### 2024-09-17 cognitive-bias.qmd
20240917 - missing cite

The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
As noted by @Kahneman, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].

>>>>

The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
As noted by @Kahneman2011, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].
